---
title: "voterfile"
format: html
editor: visual
---

Necessary packages

```{r}
library(tidyverse)
library(tidycensus)
library(shiny)
library(tigris)
```

Importing data, initial cleaning (data comes from <https://redistrictingdatahub.org/>)

```{r}

path <- "./data/CA_l2_2022stats_2020block.csv"

df <- read_csv(path) |> 
  mutate(
    has_block = !str_detect(geoid20, "NO"),
    state = if_else(has_block, str_sub(geoid20, 1, 2), "06"), 
    county = if_else(has_block, str_sub(geoid20, 1, 5), str_c("06", str_sub(geoid20, 1, 3))),
    tract = if_else(has_block, str_sub(geoid20, 1, 11), NA_character_),
    block_group = if_else(has_block, str_sub(geoid20, 1, 12), NA_character_)
  )
  
```

This is just a little bit of exploration of the data

```{r}
df |> 
  filter(has_block) |> 
  select(county, block_group, total_reg, starts_with("age")) |> 
  pivot_longer(
    cols = starts_with("age"),
    names_to = "age_range",
    names_prefix = "age_",
    values_to = "age_n"
  ) |> 
  group_by(block_group) |> 
  summarise(total_reg = sum(total_reg), age_range, age_n) |> 
  mutate(age_perc = age_n / total_reg) |> 
  summarise(sum(age_perc))

```

Pulling in some census data from tigris

```{r}

california_counties <- counties("California", year = 2020)

california_block_groups <- block_groups("California", year = 2020)

california_tracts <- tracts("California", year = 2020)
```

The next few blocks are joining the tigris census data (which includes plottable multipolygons) with the voterfile data

```{r}
california_counties <- california_counties |> 
  select(GEOID, NAME, geometry)

counties_list <- california_counties |> 
  select(NAME) |> 
  as.list()

california_block_groups <- california_block_groups |> 
  select(GEOID, geometry)

california_tracts <- california_tracts |> 
  select(GEOID, geometry)
```

```{r}
voterfile_block_groups <- df |> 
  filter(has_block) |> 
  rename(GEOID = block_group)
```

```{r}
voterfile_counties <- df |> 
  filter(!has_block) |> 
  rename(GEOID = county)
```

```{r}
voterfile_tracts <- df |> 
  filter(has_block) |> 
  rename(GEOID = tract)
```

```{r}
voterfile_block_groups <- voterfile_block_groups |>
  left_join(california_block_groups, by = "GEOID") |> 
  left_join(california_counties, join_by(county == GEOID)) |> 
  rename(county_name = NAME) |>
  rename(geometry = geometry.x) |> 
  select(!geometry.y)

voterfile_counties <- voterfile_counties |> 
  left_join(california_counties, by = "GEOID")

voterfile_tracts <-  voterfile_tracts |> 
  left_join(california_tracts, by = "GEOID") |> 
  left_join(california_counties, join_by(county == GEOID)) |> 
  rename(county_name = NAME) |>
  rename(geometry = geometry.x) |> 
  select(!geometry.y)
```

Visualizing the count of republican voters in San Francisco's census block groups

```{r}
voterfile_block_groups |> 
  filter(county_name == "San Francisco") |> 
  mutate(prop_rep = party_rep/total_reg) |> 
  ggplot(aes(fill = log10(party_rep))) + 
  geom_sf(aes(geometry = geometry)) +
  scale_fill_viridis_c(option = "magma") +
  xlim(122.55, 122.35)
```
